---
title: "NRSA-CA"
author: "David Murray-Stoker"
date: "11 November 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Predictor Variable Summaries

Summary statistics for all predictor variables used in the BRT analyses. Statistics include: N, mean, standard deviation, standard error, and 95% confidence intervals.

```{r Load Packages & Data, include=FALSE}
library(plyr) # Load before dplyr to avoid any potential issues with both packages

library(dplyr)
library(kableExtra)
library(knitr)
library(reshape2)

## Read in workspace
load("~/Documents/Research/NRSA-CA/R Code/1-Primary Analysis/DMS-NRSA-CA-Workspace.RData")
```


## Summary of Predictors for the CUS

Calculating means and measures of variation (e.g. standard errors, confidence intervals) for predictor variables at the level of the CUS.

```{r Summaries for the CUS, include=FALSE}
final.data$CUS <- "CUS"

## Environmental----------------------------------------------------------#
NO3.CUS.summary <- ddply(final.data, c("CUS"), summarise,
										     N    = sum(!is.na(NO3)),
										     mean = mean(NO3, na.rm = TRUE),
										     sd   = sd(NO3, na.rm = TRUE),
										     se   = sd / sqrt(N),
										     LL   = quantile(NO3, 0.025),
										     UL   = quantile(NO3, 0.975)
)

NH4.CUS.summary <- ddply(final.data, c("CUS"), summarise,
										     N    = sum(!is.na(NH4)),
										     mean = mean(NH4, na.rm = TRUE),
										     sd   = sd(NH4, na.rm = TRUE),
										     se   = sd / sqrt(N),
										     LL   = quantile(NH4, 0.025),
										     UL   = quantile(NH4, 0.975)
)

total.P.CUS.summary <- ddply(final.data, c("CUS"), summarise,
										         N    = sum(!is.na(total.P)),
										         mean = mean(total.P, na.rm = TRUE),
										         sd   = sd(total.P, na.rm = TRUE),
										         se   = sd / sqrt(N),
										         LL   = quantile(total.P, 0.025),
										         UL   = quantile(total.P, 0.975)
)

pH.lab.CUS.summary <- ddply(final.data, c("CUS"), summarise,
											      N    = sum(!is.na(pH.lab)),
											      mean = mean(pH.lab, na.rm = TRUE),
											      sd   = sd(pH.lab, na.rm = TRUE),
											      se   = sd / sqrt(N),
											      LL   = quantile(pH.lab, 0.025),
											      UL   = quantile(pH.lab, 0.975)
)

cond.CUS.summary <- ddply(final.data, c("CUS"), summarise,
										      N    = sum(!is.na(cond)),
										      mean = mean(cond, na.rm = TRUE),
										      sd   = sd(cond, na.rm = TRUE),
										      se   = sd / sqrt(N),
										      LL   = quantile(cond, 0.025),
										      UL   = quantile(cond, 0.975)
)

DOC.CUS.summary <- ddply(final.data, c("CUS"), summarise,
										     N    = sum(!is.na(DOC)),
										     mean = mean(DOC, na.rm = TRUE),
										     sd   = sd(DOC, na.rm = TRUE),
										     se   = sd / sqrt(N),
										     LL   = quantile(DOC, 0.025),
										     UL   = quantile(DOC, 0.975)
)

LWD.reach.CUS.summary <- ddply(final.data, c("CUS"), summarise,
													     N    = sum(!is.na(LWD.reach)),
										 	         mean = mean(LWD.reach, na.rm = TRUE),
											         sd   = sd(LWD.reach, na.rm = TRUE),
											         se   = sd / sqrt(N),
											         LL   = quantile(LWD.reach, 0.025),
											         UL   = quantile(LWD.reach, 0.975)
)

NAT.cover.CUS.summary <- ddply(final.data, c("CUS"), summarise,
											         N    = sum(!is.na(NAT.cover)),
											         mean = mean(NAT.cover, na.rm = TRUE),
											         sd   = sd(NAT.cover, na.rm = TRUE),
											         se   = sd / sqrt(N),
											         LL   = quantile(NAT.cover, 0.025),
											         UL   = quantile(NAT.cover, 0.975)
)

ALG.cover.CUS.summary <- ddply(final.data, c("CUS"), summarise,
												       N    = sum(!is.na(ALG.cover)),
												       mean = mean(ALG.cover, na.rm = TRUE),
												       sd   = sd(ALG.cover, na.rm = TRUE),
												       se   = sd / sqrt(N),
												       LL   = quantile(ALG.cover, 0.025),
												       UL   = quantile(ALG.cover, 0.975)
)

## Landscape--------------------------------------------------------------#
pct.for.CUS.summary <- ddply(final.data, c("CUS"), summarise,
												     N    = sum(!is.na(pct.for)),
												     mean = mean(pct.for, na.rm = TRUE),
												     sd   = sd(pct.for, na.rm = TRUE),
												     se   = sd / sqrt(N),
												     LL   = quantile(pct.for, 0.025),
												     UL   = quantile(pct.for, 0.975)
)

pct.ag.CUS.summary <- ddply(final.data, c("CUS"), summarise,
											      N    = sum(!is.na(pct.ag)),
											      mean = mean(pct.ag, na.rm = TRUE),
											      sd   = sd(pct.ag, na.rm = TRUE),
											      se   = sd / sqrt(N),
											      LL   = quantile(pct.ag, 0.025),
											      UL   = quantile(pct.ag, 0.975)
)

pct.urb.CUS.summary <- ddply(final.data, c("CUS"), summarise,
											       N    = sum(!is.na(pct.urb)),
											       mean = mean(pct.urb, na.rm = TRUE),
											       sd   = sd(pct.urb, na.rm = TRUE),
											       se   = sd / sqrt(N),
											       LL   = quantile(pct.urb, 0.025),
											       UL   = quantile(pct.urb, 0.975)
)

pct.ISC.CUS.summary <- ddply(final.data, c("CUS"), summarise,
											       N    = sum(!is.na(pct.ISC)),
											       mean = mean(pct.ISC, na.rm = TRUE),
											       sd   = sd(pct.ISC, na.rm = TRUE),
											       se   = sd / sqrt(N),
											       LL   = quantile(pct.ISC, 0.025),
											       UL   = quantile(pct.ISC, 0.975)
)

## Network----------------------------------------------------------------#
lat.CUS.summary <- ddply(final.data, c("CUS"), summarise,
										     N    = sum(!is.na(site.lat)),
										     mean = mean(site.lat, na.rm = TRUE),
										     sd   = sd(site.lat, na.rm = TRUE),
										     se   = sd / sqrt(N),
										     LL   = quantile(site.lat, 0.025),
										     UL   = quantile(site.lat, 0.975)
)

long.CUS.summary <- ddply(final.data, c("CUS"), summarise,
										      N    = sum(!is.na(site.long)),
										      mean = mean(site.long, na.rm = TRUE),
										      sd   = sd(site.long, na.rm = TRUE),
										      se   = sd / sqrt(N),
										      LL   = quantile(site.long, 0.025),
										      UL   = quantile(site.long, 0.975)
)

basin.area.CUS.summary <- ddply(final.data, c("CUS"), summarise,
														    N    = sum(!is.na(basin.area)),
														    mean = mean(basin.area, na.rm = TRUE),
														    sd   = sd(basin.area, na.rm = TRUE),
														    se   = sd / sqrt(N),
														    LL   = quantile(basin.area, 0.025),
														    UL   = quantile(basin.area, 0.975)
)

mean.annual.flow.CUS.summary <- ddply(final.data, c("CUS"), summarise,
													            N    = sum(!is.na(mean.annual.flow)),
													            mean = mean(mean.annual.flow, na.rm = TRUE),
													            sd   = sd(mean.annual.flow, na.rm = TRUE),
													            se   = sd / sqrt(N),
													            LL   = quantile(mean.annual.flow, 0.025),
													            UL   = quantile(mean.annual.flow, 0.975)
)

site.centrality.CUS.summary <- ddply(final.data, c("CUS"), summarise,
														         N    = sum(!is.na(site.centrality)),
														         mean = mean(site.centrality, na.rm = TRUE),
														         sd   = sd(site.centrality, na.rm = TRUE),
														         se   = sd / sqrt(N),
														         LL   = quantile(site.centrality, 0.025),
														         UL   = quantile(site.centrality, 0.975)
)
```


```{r NO3 Summary-CUS, echo=FALSE}
kable(NO3.CUS.summary, booktabs = TRUE, caption = "Summary of NO3 for the CUS.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r NH4 Summary-CUS, echo=FALSE}
kable(NH4.CUS.summary, booktabs = TRUE, caption = "Summary of NH4 for the CUS.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Total P Summary-CUS, echo=FALSE}
kable(total.P.CUS.summary, booktabs = TRUE, caption = "Summary of total P for the CUS.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r pH Summary-CUS, echo=FALSE}
kable(pH.lab.CUS.summary, booktabs = TRUE, caption = "Summary of pH for the CUS.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Conductivity Summary-CUS, echo=FALSE}
kable(cond.CUS.summary, booktabs = TRUE, caption = "Summary of conductivity for the CUS.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r DOC Summary-CUS, echo=FALSE}
kable(DOC.CUS.summary, booktabs = TRUE, caption = "Summary of DOC for the CUS.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r LWD Summary-CUS, echo=FALSE}
kable(LWD.reach.CUS.summary, booktabs = TRUE, caption = "Summary of LWD for the CUS.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Natural Cover Summary-CUS, echo=FALSE}
kable(NAT.cover.CUS.summary, booktabs = TRUE, caption = "Summary of natural cover for the CUS.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Algal Cover Summary-CUS, echo=FALSE}
kable(ALG.cover.CUS.summary, booktabs = TRUE, caption = "Summary of algal cover for the CUS.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Percent Forested Land Cover Summary-CUS, echo=FALSE}
kable(pct.for.CUS.summary, booktabs = TRUE, caption = "Summary of percent forested land cover for the CUS.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Percent Agricultural Land Cover Summary-CUS, echo=FALSE}
kable(pct.ag.CUS.summary, booktabs = TRUE, caption = "Summary of percent agricultural land cover for the CUS.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Percent Urban Land Cover Summary-CUS, echo=FALSE}
kable(pct.urb.CUS.summary, booktabs = TRUE, caption = "Summary of percent urban land cover for the CUS.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Percent Impervious Land Cover Summary-CUS, echo=FALSE}
kable(pct.ISC.CUS.summary, booktabs = TRUE, caption = "Summary of percent impervious land cover for the CUS.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Latitude Summary-CUS, echo=FALSE}
kable(lat.CUS.summary, booktabs = TRUE, caption = "Summary of site latitude for the CUS.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Longitude Summary-CUS, echo=FALSE}
kable(long.CUS.summary, booktabs = TRUE, caption = "Summary of site longitude for the CUS.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Basin Area Summary-CUS, echo=FALSE}
kable(pct.for.CUS.summary, booktabs = TRUE, caption = "Summary of basin area for the CUS.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Mean Annual Flow Summary-CUS, echo=FALSE}
kable(mean.annual.flow.CUS.summary, booktabs = TRUE, digits = 15, caption = "Summary of mean annual flow for the CUS.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Site Centrality Summary-CUS, echo=FALSE}
kable(site.centrality.CUS.summary, booktabs = TRUE, caption = "Summary of site centrality for the CUS.") %>% kable_styling(latex_options = "HOLD_position")
```


\newpage
## Summary of Predictors for the Ecoregions

Calculating means and measures of variation (e.g. standard errors, confidence intervals) for predictor variables for each of the nine ecoregions.

```{r Summaries for each Ecoregion, include=FALSE}
## Environmental----------------------------------------------------------#
NO3.summary <- ddply(final.data, c("ecoregion"), summarise,
										 N    = sum(!is.na(NO3)),
										 mean = mean(NO3, na.rm = TRUE),
										 sd   = sd(NO3, na.rm = TRUE),
										 se   = sd / sqrt(N),
										 LL   = quantile(NO3, 0.025),
										 UL   = quantile(NO3, 0.975)
)

NH4.summary <- ddply(final.data, c("ecoregion"), summarise,
										 N    = sum(!is.na(NH4)),
										 mean = mean(NH4, na.rm = TRUE),
										 sd   = sd(NH4, na.rm = TRUE),
										 se   = sd / sqrt(N),
										 LL   = quantile(NH4, 0.025),
										 UL   = quantile(NH4, 0.975)
)

total.P.summary <- ddply(final.data, c("ecoregion"), summarise,
										     N    = sum(!is.na(total.P)),
										     mean = mean(total.P, na.rm = TRUE),
										     sd   = sd(total.P, na.rm = TRUE),
										     se   = sd / sqrt(N),
										     LL   = quantile(total.P, 0.025),
										     UL   = quantile(total.P, 0.975)
)

pH.lab.summary <- ddply(final.data, c("ecoregion"), summarise,
											  N    = sum(!is.na(pH.lab)),
											  mean = mean(pH.lab, na.rm = TRUE),
											  sd   = sd(pH.lab, na.rm = TRUE),
											  se   = sd / sqrt(N),
											  LL   = quantile(pH.lab, 0.025),
											  UL   = quantile(pH.lab, 0.975)
)

cond.summary <- ddply(final.data, c("ecoregion"), summarise,
										  N    = sum(!is.na(cond)),
										  mean = mean(cond, na.rm = TRUE),
										  sd   = sd(cond, na.rm = TRUE),
										  se   = sd / sqrt(N),
										  LL   = quantile(cond, 0.025),
										  UL   = quantile(cond, 0.975)
)

DOC.summary <- ddply(final.data, c("ecoregion"), summarise,
										 N    = sum(!is.na(DOC)),
										 mean = mean(DOC, na.rm = TRUE),
										 sd   = sd(DOC, na.rm = TRUE),
										 se   = sd / sqrt(N),
										 LL   = quantile(DOC, 0.025),
										 UL   = quantile(DOC, 0.975)
)

LWD.reach.summary <- ddply(final.data, c("ecoregion"), summarise,
											     N    = sum(!is.na(LWD.reach)),
											     mean = mean(LWD.reach, na.rm = TRUE),
											     sd   = sd(LWD.reach, na.rm = TRUE),
											     se   = sd / sqrt(N),
											     LL   = quantile(LWD.reach, 0.025),
											     UL   = quantile(LWD.reach, 0.975)
)

NAT.cover.summary <- ddply(final.data, c("ecoregion"), summarise,
											     N    = sum(!is.na(NAT.cover)),
											     mean = mean(NAT.cover, na.rm = TRUE),
											     sd   = sd(NAT.cover, na.rm = TRUE),
											     se   = sd / sqrt(N),
											     LL   = quantile(NAT.cover, 0.025),
											     UL   = quantile(NAT.cover, 0.975)
)

ALG.cover.summary <- ddply(final.data, c("ecoregion"), summarise,
												   N    = sum(!is.na(ALG.cover)),
												   mean = mean(ALG.cover, na.rm = TRUE),
												   sd   = sd(ALG.cover, na.rm = TRUE),
												   se   = sd / sqrt(N),
												   LL   = quantile(ALG.cover, 0.025),
												   UL   = quantile(ALG.cover, 0.975)
)

## Landscape--------------------------------------------------------------#
pct.for.summary <- ddply(final.data, c("ecoregion"), summarise,
												 N    = sum(!is.na(pct.for)),
												 mean = mean(pct.for, na.rm = TRUE),
												 sd   = sd(pct.for, na.rm = TRUE),
												 se   = sd / sqrt(N),
												 LL   = quantile(pct.for, 0.025),
												 UL   = quantile(pct.for, 0.975)
)

pct.ag.summary <- ddply(final.data, c("ecoregion"), summarise,
											  N    = sum(!is.na(pct.ag)),
											  mean = mean(pct.ag, na.rm = TRUE),
											  sd   = sd(pct.ag, na.rm = TRUE),
											  se   = sd / sqrt(N),
											  LL   = quantile(pct.ag, 0.025),
											  UL   = quantile(pct.ag, 0.975)
)

pct.urb.summary <- ddply(final.data, c("ecoregion"), summarise,
											   N    = sum(!is.na(pct.urb)),
											   mean = mean(pct.urb, na.rm = TRUE),
											   sd   = sd(pct.urb, na.rm = TRUE),
											   se   = sd / sqrt(N),
											   LL   = quantile(pct.urb, 0.025),
											   UL   = quantile(pct.urb, 0.975)
)

pct.ISC.summary <- ddply(final.data, c("ecoregion"), summarise,
											   N    = sum(!is.na(pct.ISC)),
											   mean = mean(pct.ISC, na.rm = TRUE),
											   sd   = sd(pct.ISC, na.rm = TRUE),
											   se   = sd / sqrt(N),
											   LL   = quantile(pct.ISC, 0.025),
											   UL   = quantile(pct.ISC, 0.975)
)

## Network----------------------------------------------------------------#
lat.summary <- ddply(final.data, c("ecoregion"), summarise,
										 N    = sum(!is.na(site.lat)),
										 mean = mean(site.lat, na.rm = TRUE),
										 sd   = sd(site.lat, na.rm = TRUE),
										 se   = sd / sqrt(N),
										 LL   = quantile(site.lat, 0.025),
										 UL   = quantile(site.lat, 0.975)
)

long.summary <- ddply(final.data, c("ecoregion"), summarise,
										  N    = sum(!is.na(site.long)),
										  mean = mean(site.long, na.rm = TRUE),
										  sd   = sd(site.long, na.rm = TRUE),
										  se   = sd / sqrt(N),
										  LL   = quantile(site.long, 0.025),
										  UL   = quantile(site.long, 0.975)
)

basin.area.summary <- ddply(final.data, c("ecoregion"), summarise,
														N    = sum(!is.na(basin.area)),
														mean = mean(basin.area, na.rm = TRUE),
														sd   = sd(basin.area, na.rm = TRUE),
														se   = sd / sqrt(N),
														LL   = quantile(basin.area, 0.025),
														UL   = quantile(basin.area, 0.975)
)

mean.annual.flow.summary <- ddply(final.data, c("ecoregion"), summarise,
														      N    = sum(!is.na(mean.annual.flow)),
														      mean = mean(mean.annual.flow, na.rm = TRUE),
														      sd   = sd(mean.annual.flow, na.rm = TRUE),
														      se   = sd / sqrt(N),
														      LL   = quantile(mean.annual.flow, 0.025),
														      UL   = quantile(mean.annual.flow, 0.975)
)

site.centrality.summary <- ddply(final.data, c("ecoregion"), summarise,
														     N    = sum(!is.na(site.centrality)),
														     mean = mean(site.centrality, na.rm = TRUE),
														     sd   = sd(site.centrality, na.rm = TRUE),
														     se   = sd / sqrt(N),
														     LL   = quantile(site.centrality, 0.025),
														     UL   = quantile(site.centrality, 0.975)
)
```

```{r NO3 Summary-Ecoregions, echo=FALSE}
kable(NO3.summary, booktabs = TRUE, caption = "Summary of NO3 for each ecoregion.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r NH4 Summary-Ecoregions, echo=FALSE}
kable(NH4.summary, booktabs = TRUE, caption = "Summary of NH4 for each ecoregion.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Total P Summary-Ecoregions, echo=FALSE}
kable(total.P.summary, booktabs = TRUE, caption = "Summary of total P for each ecoregion.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r pH Summary-Ecoregions, echo=FALSE}
kable(pH.lab.summary, booktabs = TRUE, caption = "Summary of pH for each ecoregion.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Conductivity Summary-Ecoregions, echo=FALSE}
kable(cond.summary, booktabs = TRUE, caption = "Summary of conductivity for each ecoregion.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r DOC Summary-Ecoregions, echo=FALSE}
kable(DOC.summary, booktabs = TRUE, caption = "Summary of DOC for each ecoregion.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r LWD Summary-Ecoregions, echo=FALSE}
kable(LWD.reach.summary, booktabs = TRUE, caption = "Summary of LWD for each ecoregion.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Natural Cover Summary-Ecoregions, echo=FALSE}
kable(NAT.cover.summary, booktabs = TRUE, caption = "Summary of natural cover for each ecoregion.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Algal Cover Summary-Ecoregions, echo=FALSE}
kable(ALG.cover.summary, booktabs = TRUE, caption = "Summary of algal cover for each ecoregion.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Percent Forested Land Cover Summary-Ecoregions, echo=FALSE}
kable(pct.for.summary, booktabs = TRUE, caption = "Summary of percent forested land cover for each ecoregion.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Percent Agricultural Land Cover Summary-Ecoregions, echo=FALSE}
kable(pct.ag.summary, booktabs = TRUE, caption = "Summary of percent agricultural land cover for each ecoregion.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Percent Urban Land Cover Summary-Ecoregions, echo=FALSE}
kable(pct.urb.summary, booktabs = TRUE, caption = "Summary of percent urban land cover for each ecoregion.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Percent Impervious Land Cover Summary-Ecoregions, echo=FALSE}
kable(pct.ISC.summary, booktabs = TRUE, caption = "Summary of percent impervious land cover for each ecoregion.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Latitude Summary-Ecoregions, echo=FALSE}
kable(lat.summary, booktabs = TRUE, caption = "Summary of site latitude for each ecoregion.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Longitude Summary-Ecoregions, echo=FALSE}
kable(long.summary, booktabs = TRUE, caption = "Summary of site longitude for each ecoregion.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Basin Area Summary-Ecoregions, echo=FALSE}
kable(basin.area.summary, booktabs = TRUE, caption = "Summary of basin area for each ecoregion.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Mean Annual Flow Summary-Ecoregions, echo=FALSE}
kable(mean.annual.flow.summary, booktabs = TRUE, digits = 15, caption = "Summary of mean annual flow for each ecoregion.") %>% kable_styling(latex_options = "HOLD_position")
```

```{r Site Centrality Summary-Ecoregions, echo=FALSE}
kable(site.centrality.summary, booktabs = TRUE, caption = "Summary of site centrality for each ecoregion.") %>% kable_styling(latex_options = "HOLD_position")
```


